Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2025-05-14 paper.qmd
Updated Figure 10

error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs, 
      aes(x = sqrt_row_wise_total_error, 
          y = density,
          colour = factor(error_cat_n, levels = c("deselected", "selected")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("b1",
                      position = c(0.95, 0.9),
                      cex = 2) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))

error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs,
        aes(x = sqrt_row_wise_total_error, 
            y = density,
            colour = factor(error_cat_n2, levels = c("deselected", "selected")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("a1",
                      position = c(0.95, 0.9),
                      cex = 2) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))
```

```{r}
#| label: link-model-one-dens-clust
plot_tsne_dens <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n2,
                             levels = c("deselected", "selected")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1) +
  interior_annotation("a2",
                      position = c(0.08, 0.9),
                      cex = 2)

plot_tsne_dens_selected <- error_df_one_curvy_abs |>
  ggplot(aes(x = tSNE1,
             y = tSNE2,
             colour = factor(error_cat_n,
                             levels = c("deselected", "selected")))) +
  geom_point(alpha=0.7) +
  scale_color_manual(values=c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1) +
  interior_annotation("b2",
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
                         lineFrom = tr_from_to_df_dens_clust$from,
                         lineTo = tr_from_to_df_dens_clust$to,
                         group = factor(df_model_data_n$type,
                                        c("deselected", "selected", "model")),
                         levelColors = c('#d9d9d9', '#800026', "#000000"))
```

@@

five_c_shaped_proj_tsne_model1 <- plot_proj(
    proj_obj = proj_obj1,
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    axis_text_size = 4.5, 
    is_category = TRUE) +
  interior_annotation(label = "a3", position = c(0.08, 0.9), cex = 2) +
  scale_color_manual(values = c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1,
        legend.position = "none")

@@

five_c_shaped_proj_tsne_model1_selected <- plot_proj(
    proj_obj = proj_obj1, 
    point_param = c(1.5, 0.5), # size, alpha, color
    line_param = c(0.8, 0.4, "#000000"), # linewidth, alpha
    plot_limits = c(-0.35, 0.35), 
    axis_text_size = 4.5, 
    is_category = TRUE) +
  interior_annotation(label = "b3", position = c(0.08, 0.9), cex = 2) +
  scale_color_manual(values = c('#d9d9d9', '#800026')) +
  theme(aspect.ratio = 1,
        legend.position = "none")
```

>>>>

error_df_one_curvy_abs_selected11 <- error_df_one_curvy_abs |>
  dplyr::filter(error_cat_n == "selected")

error_df_one_curvy_abs_deselected11 <- error_df_one_curvy_abs |>
  dplyr::filter(error_cat_n == "deselected")

error_plot_one_curvy_hist_selected <- ggplot(error_df_one_curvy_abs_deselected11, 
      aes(x = sqrt_row_wise_total_error, 
          y = density)) +
  geom_point(alpha=0.7, color = "#d9d9d9") +
  geom_point(data = error_df_one_curvy_abs_selected11, 
      aes(x = sqrt_row_wise_total_error, 
          y = density),
      alpha=0.7, color = "#800026", size = 3) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("b1",
                      position = c(0.95, 0.9),
                      cex = 2) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))

error_df_one_curvy_abs_selected12 <- error_df_one_curvy_abs |>
  dplyr::filter(error_cat_n2 == "selected")

error_df_one_curvy_abs_deselected12 <- error_df_one_curvy_abs |>
  dplyr::filter(error_cat_n2 == "deselected")

error_plot_one_curvy_hist <- ggplot(error_df_one_curvy_abs_deselected12, 
      aes(x = sqrt_row_wise_total_error, 
          y = density)) +
  geom_point(alpha=0.7, color = "#d9d9d9") +
  geom_point(data = error_df_one_curvy_abs_selected12, 
      aes(x = sqrt_row_wise_total_error, 
          y = density),
      alpha=0.7, color = "#800026", size = 3) +
  xlab(expression(e[hj])) +
  ylab("") +
  interior_annotation("a1",
                      position = c(0.95, 0.9),
                      cex = 2) +
  theme_bw() +
  theme(aspect.ratio = 1,
        axis.title = element_text(size = 18),
        axis.text = element_text(size = 15))
```

```{r}
#| label: link-model-one-dens-clust

plot_tsne_dens <- error_df_one_curvy_abs_deselected12 |>
  ggplot(aes(x = tSNE1,
             y = tSNE2)) +
  geom_point(alpha=0.7, color = "#d9d9d9") +
  geom_point(data = error_df_one_curvy_abs_selected12,
             aes(x = tSNE1, y = tSNE2),
             alpha=0.7, color = "#800026", size = 3) +
  theme(aspect.ratio = 1) +
  interior_annotation("a2",
                      position = c(0.08, 0.9),
                      cex = 2)

plot_tsne_dens_selected <- error_df_one_curvy_abs_deselected11 |>
  ggplot(aes(x = tSNE1,
             y = tSNE2)) +
  geom_point(alpha=0.7, color = "#d9d9d9") +
  geom_point(data = error_df_one_curvy_abs_selected11,
             aes(x = tSNE1, y = tSNE2),
             alpha=0.7, color = "#800026", size = 3) +
  theme(aspect.ratio = 1) +
  interior_annotation("b2",
                      position = c(0.08, 0.9),
                      cex = 2)
```

@@

langevitour::langevitour(df_model_data_n[1:(length(df_model_data_n)-1)],
                         lineFrom = tr_from_to_df_dens_clust$from,
                         lineTo = tr_from_to_df_dens_clust$to,
                         group = factor(df_model_data_n$type,
                                        c("deselected", "selected", "model")),
                         levelColors = c('#d9d9d9', '#800026', "#000000"),
                         pointSize = c(1.5, 3, 0))
```

@@

projected_df <- projected_df |>
      dplyr::mutate(cluster = proj_obj1$cluster)

five_c_shaped_proj_tsne_model1 <- ggplot() +
      geom_point(
        data = projected_df,
        aes(
          x = proj1,
          y = proj2,
          colour = factor(cluster, levels = c("deselected", "selected")),
          size = factor(cluster, levels = c("deselected", "selected"))),
        alpha = 0.5) +
      geom_segment(
        data = model_df,
        aes(
          x = proj1_from,
          y = proj2_from,
          xend = proj1_to,
          yend = proj2_to),
        color = "#000000",
        linewidth = 0.8,
        alpha = 0.4) +
      geom_segment(
          data=axes,
          aes(x=x1, y=y1, xend=x2, yend=y2),
          colour="grey70") +
        geom_text(
          data=axes,
          aes(x=x2, y=y2),
          label=rownames(axes),
          colour="grey50",
          size = 4.5) +
        geom_path(
          data=circle,
          aes(x=c1, y=c2), colour="grey70") +
        xlim(c(-0.35, 0.35)) +
        ylim(c(-0.35, 0.35)) +
        interior_annotation(label = "a3", position = c(0.08, 0.9), cex = 2) +
        scale_color_manual(values = c('#d9d9d9', '#800026')) +
        scale_size_manual(values = c('deselected' = 1.5, 'selected' = 3)) +
        theme(aspect.ratio = 1,
              legend.position = "none")

@@

projected_df <- projected_df |>
      dplyr::mutate(cluster = proj_obj1$cluster)

five_c_shaped_proj_tsne_model1_selected <- ggplot() +
      geom_point(
        data = projected_df,
        aes(
          x = proj1,
          y = proj2,
          colour = factor(cluster, levels = c("deselected", "selected")),
          size = factor(cluster, levels = c("deselected", "selected"))),
        alpha = 0.5) +
      geom_segment(
        data = model_df,
        aes(
          x = proj1_from,
          y = proj2_from,
          xend = proj1_to,
          yend = proj2_to),
        color = "#000000",
        linewidth = 0.8,
        alpha = 0.4) +
      geom_segment(
          data=axes,
          aes(x=x1, y=y1, xend=x2, yend=y2),
          colour="grey70") +
        geom_text(
          data=axes,
          aes(x=x2, y=y2),
          label=rownames(axes),
          colour="grey50",
          size = 4.5) +
        geom_path(
          data=circle,
          aes(x=c1, y=c2), colour="grey70") +
        xlim(c(-0.35, 0.35)) +
        ylim(c(-0.35, 0.35)) +
  interior_annotation(label = "b3", position = c(0.08, 0.9), cex = 2) +
  scale_color_manual(values = c('#d9d9d9', '#800026')) +
  scale_size_manual(values = c('deselected' = 1.5, 'selected' = 3)) +
  theme(aspect.ratio = 1,
        legend.position = "none")
```
